Lens Modeling in \sw

### 2014-06-05 text/partA.tex
edits

@@

For the sake of blind testing, the information in this section was
revealed neither to RK, while choosing the challenge set of 29 sims,
nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
modeler' PS) until the modeling stage was done.

@@

@@

@@

% Rafi's original one-coumn design:
%
% \begin{figure}
%   \centering
%   \subfigure[real mass distribution]{
%     \label{fig:6941_sim_mass}
%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
%   }
%   \subfigure[real arrival-time surface]{
%     \label{fig:6941_sim_arr}
%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
%   }
%   \subfigure[model mass distribution]{
%     \label{fig:6941_mass}
%     \includegraphics[width=0.45\textwidth]{fig/006941_mass}
%   }
%   \subfigure[model arrival-time surface]{
%     \label{fig:6941_cont}
%     \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
%   }
%   \subfigure[real vs model enclosed mass]{
%     \label{fig:6941_kappa}
%     \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
%   }
%   \subfigure[model lensed image]{
%     \label{fig:6941_atime}
%     \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
%   }
%   \caption[result 6941 (ASW000102p)]{Model results for a simple
%     simulated lens.  (The label ASW000102p in the lower-left panel is
%     the \sw identifier for the sky patch where this sim was placed.)
%     The model input is shown in the top panel of
%     Figure~\ref{fig:input-spag}.  For details of the individual panels
%     here, see section \ref{sec:example_models}.}
%   \label{fig:6941}
% \end{figure}

% Rafik's two row design

\newpage

@@

\newpage

@@

Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
saddle points are correctly identified, and the orientation of the
ellipticity of the mass distribution is correctly reproduced.  The
Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
shows another model of the same system.  In this one, the
identification of the minima and saddle points was incorrect, and mass
distribution comes out elongated East-West instead of North-South.
The mass distribution also appears somewhat jagged and the
saddle-point contours are not as clean as in the previous examples;
these are often indicators of a problem with the model.  The enclosed
mass is, however, none the worse --- the reason is probably that in a
relatively symmetrical image configuration, the Einstein radius is quite
well constrained by the images in a fairly model-independent way.

>>>>

A challenge set of 29 was selected, representing the different image
morphologies among the spacewarps sims.  Modellers then contributed
129 models for these sims (at least for each sim).  Models were
collecting in the same forum used to model real candidate lenses, and
under the same conditions --- modellers were free to consult and
refine each other's models, but had no information on how the sims
were generated.

@@

@@

The information in this section was not revealed to the person
choosing the challenge set (RK) or to the modellers (EB, CC, CM, PS,
JO and JW) until the modeling stage was done.

@@

The community of modellers proposed a total of 129 models for the 29 sims in the challenge. Typically XX models were produced per simulation, with XX systems having only XX models (corresponding tot he most simplest simulations) to XX having XX models (a complex and contrversial lens system). In the following we discuss eight example lenses in detail with individual notes per system and corresponding plots are shown in figures 4-11 (or whatever compact version you end up making). These eight examples were chosen on the basis of …??

@@

\vfill\penalty-2000

@@

\vfill\penalty-2000

@@

Figure \ref{fig:7022} shows another model of the same system.  In this
one, the identification of the minima and saddle points was incorrect,
and mass distribution comes out elongated East-West instead of
North-South.  The mass distribution also appears somewhat jagged and
the saddle-point contours are not as clean as in the previous
examples; these are often indicators of a problem with the model.  The
enclosed mass is, however, none the worse --- the reason is probably
that in a relatively symmetrical image configuration, the Einstein
radius is quite well constrained by the images in a fairly
model-independent way.

Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
saddle points are correctly identified, and the orientation of the
ellipticity of the mass distribution is correctly reproduced.  The
Einstein radius is somewhat overestimated.

### 2014-06-05 text/partC.tex
edits

The middle line of equation \eqref{cfBN86}, though very different in
form to the expression \eqref{eq:tgrav} for the gravitational time
delay, is in fact equivalent.  This can be shown by considering a
multipole expansions for the solution of the latter equation.  Another
way to see the equivalence is to write a discrete version of
\eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
\begin{equation} \label{eq:relax}
\begin{aligned}
\tgrav(x,y) &=
\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
\end{aligned}
\end{equation}
The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
been replaced by the average over four neighbouring points, and the
enclosed mass $M_{\textstyle\bullet}$ has been replaced by
$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
recognisable as a formula for solving the two-dimensional Poisson
equation from \eqref{cfBN86} by relaxation.

>>>>

The middle line of equation \eqref{cfBN86}, though very different in
form to the expression \eqref{eq:tgrav} for the gravitational time
delay, is in fact equivalent.  This can be shown by considering a
multipole expansions for the solution of the latter equation.  Another
way to see the equivalence is to write a discrete version of
\eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
\begin{equation} \label{eq:relax}
\begin{aligned}
\tgrav(x,y) &=
\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
\end{aligned}
\end{equation}
The circular average
$\left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle$ has
been replaced by the average over four neighbouring points, and the
enclosed mass $M(x_\subbullet,y_\subbullet)$ has been replaced by
$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
recognisable as a formula for solving the two-dimensional Poisson
equation from \eqref{cfBN86} by relaxation.
